\simname 2.0 User's Guide

### 2009-03-16 doc/paper.tex
sitll writing the discussion section
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/tags/pbooksim@1152 ce7aba0d-9f81-40a6-a495-cecae84b0515

\subsection{Analyzing Synchronization Mechanisms}
The first step in analyzing our parallelized network simulator was to
determine the performance of the different synchronization mechanisms that
we had employed at different points in development.  Figure \ref{synch}
shows the performance of the various synchronization mechansims on a
single node of the Hotbox cluster that consists of two Intel Xeon Quad
cores.  We only ran experiments up to 8 threads as this is the maximum
number of hardware thread contexts available without encountering the
overhead of using ethernet to communicate between nodes.\\
\begin{center}
\includegraphics[width=2in,height=2in]{synch.eps} \\
\small{Figure 1: This figure shows the performance of the different
synchronization mechanisms on up to eight threads.}
\end{center}
From these results we can see that our hypotheses were confirmed regarding
the performance of the various synchronization mechanisms.  Barriering
every cycle added significant overhead to performance and ultimately
prevented even 8 threads from extracting much parallelism at all.
Surprisingly the same result was discovered when the barrier was removed and
all the queues were replaced with lock-free queues.  However, once we were
able to only implement the shared channels with lock-free queues , we could see a noticeable performance gain and even some
additional scalability.  Finally, once we fused several of the loops being
executed within the program, we were able to improve some of the temporal
locality of data use and thereby achieve a little additional
performance.\\
~\\
Despite our efforts however, we were only able to extract a two fold
speedup using eight threads, far below a linear speedup.  This was very
perplexing to us as we had consciously attempted to break up the work into
parallel computations that required very little inter-thread
communication.  In order to determine exactly what was causing us to lose
performance we then proceeded to run several experiments to find the
bottlenecks in the program.

\subsection{Determining Scalability Bottlenecks}
Having observed the somewhat poor performance of our application we then
set about attempting to determine the reason for such poor performance.
Our initial hypothesis was that application was memory bound based on the
fact that there is a significant amount of data being processed in our
simulator.  In addition to this, the fact that Pthreads is a shared memory
programming model implies that if the working set cannot fit inside the L2
cache, then there will be a significant amount of data being moved both to
and from main memory.\\
~\\
To test our hypothesis that the simulator was memory bound, we ran the
simulator again on the Hotbox cluster, but this time ran each thread on a
different node.  This allowed each thread to have access to the entire
memory bandwidth of its node, and in effect, thereby increase the total
amount of memory bandwidth available to the entire application.  The
results of this experiment can be seen in Figure \ref{smp}.\\
\begin{center}
\includegraphics[height=2in,width=2in]{smp.eps} \\
\small{Figure 2: This figure shows the performance of the simulator on
up to 16 threads on a simulated SMP machine.  The speedup is relative to
the sequential simulator.}
\end{center}
From the result of this experiment we were able to see that bandwidth was
not the primary issue.  This is due to the fact that the performance of
the simulator remained poor and was even worse than the CMP version up to
16 threads.  In addition to revealing that the poor performance was not
due to memory bandwidth, it also revealed that communication was playing a
large role in the performance as the increased cost of communicating via
ethernet between nodes caused performance to be even worse than the CMP
configuration.\\
~\\
After realizing that our program was not memory bound, we wanted to see if
the size of the problem had any effect on the speedup.  We tested this by
plotting the speedup of a four threaded run on a given network size over the
sequential time for simulating the same size network on the Hotbox
cluster.  The results can be seen in Figure \ref{probsize}.
\begin{center}
\includegraphics[height=2in,width=2in]{probsize.eps}
\small{Figure 3: This figure shows the speedup versus network size of a
four threaded run of the parallel simulator on the Hotbox cluster.}
\end{center}
From this figure we could see that essentially the speedup was independent
of problem size.  This meant that regardless of the size of the network
simulated, the four-threaded version ultimately was constrained by the
same relative to the sequential version.  The fact that the speedup
attained was significantly less than a linear speedup of 4 indicated that
the threads were not compute bound.  Instead we hypothesized that false
sharing was causing the four threads to thrash each others' caches.\\
~\\
To confirm our diagnosis of false sharing, we decided to run the simulator
on the both the Niagara T2 and T2+ machines.  Since these machines were
better designed to deal with long latency loads and stores, our hope was
that they would be able to mitigate the effects of false sharing to some
extend.  The results from this experiment can be seen in figure
\ref{niagara}.
\begin{center}
\includegraphics[height=2in,width=2in]{niagara.eps}
\small{Figure 4: This figure shows the speedup of the parallel simulator
up to 64 threads on the Niagara T2 and T2+ machines.}
\end{center}
We noticed that in this experiment, both the T2 and T2+ were able to
achieve some speedup to a certain number of threads, but then performance
began to fall off.  We determined that once the threads' working set sizes
began to interfere with each other in the shared L2 cache is when the
performance begins to degrade and couldn't offer us much insight into  any potential false sharing that was occurring when run on other
hardware platforms.

@@

\section{Conclusion \label{conc}}
We have developed a parallelized version of a interconnect network
simulator that is capable of achieving some speedup on both CMP and SMP
machines.  We outlined the basics of the simulator and the various
techniques that we used in order to extract concurrency from the
simulator.  We then demonstrated the performance of our parallel simulator
on several different hardware platforms.  Investigtions showed that the
simulator is prone to sufferring from false sharing and extreme
communication overheads.  Despite these bottlenecks we were able to
achieve up to a 3X performance which can significantly lower the
simulation time required for running future experiments.

>>>>

\subsection{Analyzing Synchronization Mechanisms}
The first step in analyzing our parallelized network simulator was to
determine the performance of the different synchronization mechanisms that
we had employed at different points in development.  Figure \ref{synch}
shows the performance of the various synchronization mechansims on a
single node of the Hotbox cluster that consists of two Intel Xeon Quad
cores.  We only ran experiments up to 8 threads as this is the maximum
number of hardware thread contexts available without encountering the
overhead of using ethernet to communicate between nodes.\\
\begin{center}
\includegraphics[width=3in,height=3in]{synch.eps} \\
\small{Figure 1: This figure shows the performance of the different
synchronization mechanisms on up to eight threads.}
\end{center}
From these results we can see that our hypotheses were confirmed regarding
the performance of the various synchronization mechanisms.  Barriering
every cycle added significant overhead to performance and ultimately
prevented even 8 threads from extracting much parallelism at all.
Surprisingly the same result was discovered when the barrier was removed and
all the queues were replaced with lock-free queues.  However, once we were
able to only implement the shared channels with lock-free queues , we could see a noticeable performance gain and even some
additional scalability.  Finally, once we fused several of the loops being
executed within the program, we were able to improve some of the temporal
locality of data use and thereby achieve a little additional
performance.\\
~\\
Despite our efforts however, we were only able to extract a two fold
speedup using eight threads, far below a linear speedup.  This was very
perplexing to us as we had consciously attempted to break up the work into
parallel computations that required very little inter-thread
communication.  In order to determine exactly what was causing us to lose
performance we then proceeded to run several experiments to find the
bottlenecks in the program.

\subsection{Determining Scalability Bottlenecks}
Having observed the somewhat poor performance of our application we then
set about attempting to determine the reason for such poor performance.
Our initial hypothesis was that application was memory bound based on the
fact that there is a significant amount of data being processed in our
simulator.  In addition to this, the fact that Pthreads is a shared memory
programming model implies that if the working set cannot fit inside the L2
cache, then there will be a significant amount of data being moved both to
and from main memory.\\
~\\
To test our hypothesis that the simulator was memory bound, we ran the
simulator again on the Hotbox cluster, but this time ran each thread on a
different node.  This allowed each thread to have access to the entire
memory bandwidth of its node, and in effect, thereby increase the total
amount of memory bandwidth available to the entire application.  The
results of this experiment can be seen in Figure \ref{smp}.\\
\begin{center}
\includegraphics[height=3in,width=3in]{smp.eps} \\
\small{Figure 2: This figure shows the performance of the simulator on
up to 16 threads on a simulated SMP machine.  The speedup is relative to
the sequential simulator.}
\end{center}
From the result of this experiment we were able to see that bandwidth was
not the primary issue.  This is due to the fact that the performance of
the simulator remained poor and was even worse than the CMP version up to
16 threads.  In addition to revealing that the poor performance was not
due to memory bandwidth, it also revealed that communication was playing a
large role in the performance as the increased cost of communicating via
ethernet between nodes caused performance to be even worse than the CMP
configuration.\\
~\\
After realizing that our program was not memory bound, we wanted to see if
the size of the problem had any effect on the speedup.  We tested this by
plotting the speedup of a four threaded run on a given network size over the
sequential time for simulating the same size network on the Hotbox
cluster.  The results can be seen in Figure \ref{probsize}.
\begin{center}
\includegraphics[height=3in,width=3in]{probsize.eps}
\small{Figure 3: This figure shows the speedup versus network size of a
four threaded run of the parallel simulator on the Hotbox cluster.}
\end{center}
From this figure we could see that essentially the speedup was independent
of problem size.  This meant that regardless of the size of the network
simulated, the four-threaded version ultimately was constrained by the
same relative to the sequential version.  The fact that the speedup
attained was significantly less than a linear speedup of 4 indicated that
the threads were not compute bound.  Instead we hypothesized that false
sharing was causing the four threads to thrash each others' caches.\\
~\\
To confirm our diagnosis of false sharing, we decided to run the simulator
on the both the Niagara T2 and T2+ machines.  Since these machines were
better designed to deal with long latency loads and stores, our hope was
that they would be able to mitigate the effects of false sharing to some
extend.  The results from this experiment can be seen in figure
\ref{niagara}.
\begin{center}
\includegraphics[height=3in,width=3in]{niagara.eps}
\small{Figure 4: This figure shows the speedup of the parallel simulator
up to 64 threads on the Niagara T2 and T2+ machines.}
\end{center}
We noticed that in this experiment, both the T2 and T2+ were able to
achieve some speedup to a certain number of threads, but then performance
began to fall off.  We determined that once the threads' working set sizes
began to interfere with each other in the shared L2 cache is when the
performance begins to degrade and couldn't offer us much insight into  any potential false sharing that was occurring when run on other
hardware platforms.

@@

\section{Conclusion \label{conc}}
We have developed a parallelized version of a interconnect network
simulator that is capable of achieving some speedup on both CMP and SMP
machines.  We outlined the basics of the simulator and the various
techniques that we used in order to extract concurrency from the
simulator.  We then demonstrated the performance of our parallel simulator
on several different hardware platforms.  Investigations showed that the
simulator is prone to suffering from false sharing and extreme
communication overheads.  Despite these bottlenecks we were able to
achieve up to a 3X performance which can significantly lower the
simulation time required for running future experiments.
